---
title: "hw03_Mangahas"
author: "Racquelle Sabrina Mangahas"
date: "27/09/2019"
output: html_document
---

```{r}
library(gapminder)
library(tidyverse)
```

```{r}
gapminder
```

```{r allow errors, echo = FALSE}
knitr::opts_chunk$set(error = TRUE, warning = FALSE)
```

# Task 1
*How is life expectancy changing over time on different continents?
*

```{r}
gapminder_LifeExp <- gapminder %>% 
  group_by(continent) %>% 
  summarise(year, lifeExp) %>% 
  DT::datatable(gapminder_LifeExp, colnames = c('Continent', 'Year', 'Life Expectancy'))
```

```{r}
gapminder_LifeExp
```


```{r, fig.width = 10, fig.height = 10}
ggplot(gapminder, aes(year, lifeExp, size=pop, fill=continent)) +
  geom_point(alpha = 0.9, shape=21) +
  facet_wrap(~continent, scales="free_x") +
  xlab("Year") +
  ylab("Life Expectancy") +
  theme_minimal() +
  labs(fill="Continent", size="Population") +
  ggtitle("Change of Life Expectancy Over Time") +
  theme(plot.title = element_text(hjust = 0.5))
```

# Task 2
*Get all the maximum and minimum of GDP per capita for all continents.

```{r, fig.width = 10, fig.height = 10}
ggplot(gapminder, aes(continent, gdpPercap)) +
  geom_boxplot(alpha=1) +
  scale_y_continuous(labels = scales::comma_format()) +
  xlab("Continent") +
  ylab("GDP per capita ($)") +
  theme_linedraw() +
  guides(fill="none") +
  ggtitle("GDP per capita for each Continent") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
gapminder_MinMax <-gapminder %>% 
  group_by(continent) %>% 
  summarise(min(gdpPercap), max(gdpPercap)) %>% 
DT::datatable(gapminder_MinMax, colnames = c('Continent', 'Minimum GDP per capita', 'Maximum GDP per capita'))
```

```{r}
gapminder_MinMax
```



jitter, show the max 
stats. summary function

jitter, show the max 
stats. summary function

timescale

make new max and min dataset
and combine
then create the graph 


# Task 3
*Look at the spread of GDP per capita within the continents 

```{r}
gapminder_GDP <-gapminder %>% 
  group_by(continent) %>% 
  summarise(gdpPercap) 
```


standard deviation
interquartile range
```{r, fig.width = 10, fig.height = 10}
ggplot(gapminder, aes(continent, gdpPercap, fill=continent)) +
  geom_violin(alpha=1) +
  scale_y_continuous(labels = scales::comma_format()) +
  xlab("Continent") +
  ylab("GDP per capita ($)") +
  theme_linedraw() +
  guides(fill="none") +
  ggtitle("Spread of GDP per capita for each Continent") +
  theme(plot.title = element_text(hjust = 0.5))
```




